Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-05-06 article.tex
Describe artifical data sets

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all HLA-A and 
most HLA-B super types (\cite{bianchi2017}). 
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid 
bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.
Note that the mechanism by which a cell presents its TMHs is
yet unknown.

@@

@@

Because we use \richel{X} MHC-I and \richel{Y} MHC-II haplotypes
for a significance level of 5\%, by chance we expect \richel{X+Y/20}
times to find a (MHC-I or MHC-II) haplotype 
that is suggested to have selection working 
upon it. Applying a binomial distribution ($p = 0.05$, $n = X+Y$, $\alpha = 0.05$),
we will claim we detected selection when we found \richel{Z} or
more examples of having a significant result.

@@

Of the complete system, we test for each of the 12 state transitions (that is,
we do not test for mutations that result in the same state).
For each transition rate, we know the expected and observed number
of transitions. If the observed number of transitions is below the expected
value, we test for finding that (low) number of transitions or less.
If the observed number of transitions is above the expected
value, we test for finding that (high) number of transitions or more.
We use a significance level $\alpha$ of 5\%, because we have no prior
evidence.

As we are testing 12 hypothesis, 
we correct against type I errors (false positives), 
by applying the Holm-Bonferroni correction, 
resulting in \richel{here is just a normal Bonferroni correction, for now}
$\alpha_c = \frac{\alpha}{12}$.
If the probability of observing that few/many 
transition is below $\alpha_c$, we will state 
that this unlikely situation may be caused by natural selection.

@@

@@

% Poisson binomial distribution

@@

We found that the percentages of epitopes overlapping 
with TMHs for a human and COVID-19 proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.

We found that the percentages of epitopes overlapping 
with TMHs for a human and Mycobacterium proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.

	% We conclude that MHC-II binds to TMH peptides with a higher/lower/equal
	% probability than expected by chance.

	% We conclude that the evolutionary conservation if the TMH parts of membrane
	% proteins is higher/less/equal compare to its non-TMH counterparts.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.
Because the full COVID-19 only has 4 TMHs \richel{check}, the percentages
of MHC-I epitopes being part of a TMH are likelier to be affected by
stochasticity. We chose to use COVID-19 regardless, as the thousands
of its time-dated genomic sequences are ideal for determining the 
evolutionary conservation of MHC-I detecting TMHs.

>>>>

For MHC-I, it was found that predicted epitopes derived 
from human transmembrane helices (TMHs)
are over-presented by all 5 HLA-A and 
most of 8 HLA-B super types (\cite{bianchi2017}).
One explanation is that the presentation of TMHs 
may have an evolutionary advantage for 
the (human) host, as TMHs have a reduced variability 
due to the functional requirement of being able to span a lipid bilayer. 
Due to this, pathogens have a lower chance to develop an escape mutation,
as many mutations will result in a disfunctional TMH.
Note that the mechanism by which a cell presents its TMHs is
yet unknown.

@@

\paragraph{Extracting polypeptides}

\richel{More detail here}.

@@

Because we use 13 MHC-I and 7 MHC-II haplotypes
for a significance level of 5\%, by chance we expect (5\% of $13 + 7$)
one times to find a (MHC-I or MHC-II) haplotype 
that is suggested to have selection working upon it.
Applying a binomial distribution ($p = 0.05$, $n = 13+7$, $\alpha = 0.05$),
we will claim we there is some form of selection when we find 3
or more examples of having a significant result.
However, because there are 2 hypotheses tested per haplotype,
we will claim there is directional selection, 
when, for a same hypothesis, we find 2 examples of having a significant result.

% > pbinom(q = 2, size = 20, prob = 0.05, lower.tail = FALSE)
% [1] 0.07548367
%
% pbinom(q = 3, size = 20, prob = 0.05, lower.tail = FALSE)
% [1] 0.01590153

@@

Of the complete system, we test for each of the 2 state transitions (that is,
we do not test for mutations that result in the same state).
For each transition rate, we know the expected and observed number
of transitions. If the observed number of transitions is below the expected
value, we test for finding that (low) number of transitions or less.
If the observed number of transitions is above the expected
value, we test for finding that (high) number of transitions or more.
We use a significance level $\alpha$ of 5\%, because we have no prior
evidence.

As we are testing 2 hypotheses, 
we correct against type I errors (false positives), 
by applying the Holm-Bonferroni correction, 
resulting in \richel{here is just a normal Bonferroni correction, for now}
$\alpha_c = \frac{\alpha}{2}$.
If the probability of observing that few/many 
transition is below $\alpha_c$, we will state 
that this unlikely situation may be caused by natural selection.

@@

\paragraph{Measuring strength}

To assess the strength of our analysis, 
we have applied it to four simulated data sets,
which are the product of using either MHC-I or MHC-II,
as well as using two different selection scenarios.

For both MHC-I and MHC-II, we use the haplotypes that have 
the lowest phenotypic frequency, which is \richel{X} for MHC-I
and DPB1*0501 (21.7\%) for MHC-II.

The two selection scenarios are either no selection
or a two-fold selection on what we would expect.
The scenario without selection, the 'null' data set, acts 
as a control, where we let the simulate observations occur by chance. 
The 'rigged' data set is a data set in which we 
let evolution bias the observations two-fold, that is, 
we let mutations that are beneficial occur twice as often as expected by chance.

All data sets consists out of $10^4$ observations.
Each observation starts with a randomly generated source polypeptide fragment,
as well a mutant. For the 'null' data set, the mutant is randomly select.
For the 'rigged' data set, beneficial mutations are twice as likely to
be selected. We define a beneficial mutation as a transition from
being detected to being undetected.

@@

@@

We found that the percentages of epitopes overlapping 
with TMHs for a human and viral proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.

We found that the percentages of epitopes overlapping 
with TMHs for a human and bacterial proteome are 
[similar/different]. In other words, the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its bacterial pathogen.

@@

We concluded that the
epitopes that MHC-I presents are [as/not as] likely 
to be derived from TMH within either a human host and its viral pathogen.
Because the full COVID-19 has 21 TMHs, the percentages
of MHC-I epitopes being part of a TMH are likelier to be affected by
stochasticity. We chose to use COVID-19 regardless, as the thousands
of its time-dated genomic sequences are ideal for determining the 
evolutionary conservation of MHC-I detecting TMHs.
